---
title: "Executive Governance Expenditures in WA"
output: html_document
---

#By Yiming Sun
I'm using data of [Expenditures by Candidates and Political Committees](https://data.wa.gov/Politics/Expenditures-by-Candidates-and-Political-Committee/tijg-9zyp), and I used the WA shapefile from Professor's github directly

```{r, echo=FALSE}
#preparation
library(ggthemes)
library(scales)
library(lubridate)
library(knitr)
library(tidyverse)
library(ggplot2)
```

The dataset contains expenditures made by Washington State Candidates and Political Committees for the last 10 years as reported to the Public Disclosure Commission. I exlpored the distribution of total expenditures paid for different zipcode WA recipients in 2016.
```{r, echo=FALSE}
#Putting in the data
token='&$$app_token=qkzOjPpAsJDHAxBVNro7U82j5'
endpoint="https://data.wa.gov/resource/exenditures_by_candidates_and_political_committees.json?"
filters='&jurisdiction_type=Statewide'
sqlQueries='&$where=election_year >= 2012'
limit='$limit=40000'

urlLink = paste0(endpoint,limit,filters,token)
library(jsonlite) 
expenditures = fromJSON(urlLink)
```


```{r, echo=FALSE}
#Keep only 4 columns of the data frame

expenditures= expenditures[,c('recipient_state','recipient_zip','amount','election_year','party')]
str(expenditures)
expenditures$amount=as.numeric(expenditures$amount)
expenditures$recipient_zip=as.numeric(expenditures$recipient_zip)
```
```{r, echo=FALSE}
#Clean the data-Keep recipient_state of "WA"" and election year of "2016""
#Also reset the dataframe
WA2016=expenditures[expenditures$recipient_state%in%'WA'&expenditures$election_year%in%'2016',]
row.names(WA2016)=NULL
```

```{r, echo=FALSE}
#drop the missing data
WA2016$recipient_zip=as.numeric(WA2016$recipient_zip)
WA2016=WA2016[complete.cases(WA2016),]
```

```{r, echo=FALSE}
#reset dataframe again
row.names(WA2016)=NULL
```
I hope to show the data in the following table in a map.
```{r, echo=FALSE}
#Create a new data frame that aggregate the expenditure amount according to zipcode
WA_ex_zip=aggregate(list(amount=WA2016$amount),list(Zip=WA2016$recipient_zip),sum)
names(WA_ex_zip)=c('recipient_zip','amount')
WA_ex_zip=WA_ex_zip[(WA_ex_zip$recipient_zip <=99403) & (WA_ex_zip$recipient_zip>=98001),]
head(WA_ex_zip)
```

```{r, echo=FALSE}
#Get the shapefile
compressedMap= "https://github.com/EvansDataScience/data/raw/master/WAzips.zip"
library(utils)
temp=tempfile()
download.file(compressedMap, temp)
unzip(temp)

(maps=list.files(pattern = 'shp'))
library(rgdal)
wazipMap <- rgdal::readOGR("SAEP_ZIP_Code_Tabulation_Areas.shp",stringsAsFactors=F)
```
And this is the map I created, we could see that not all the locations in WA have recipients of govenment expenditures-only pllaces with color have. To further understanding the map, the places with color blue have the highest amount of govenment expenditures, and the places with color yellow have the lowest.
```{r, echo=FALSE}
#merge data sets and shapefile
expenWAmap=merge(wazipMap,WA_ex_zip, by.x='ZCTA5CE10', by.y='recipient_zip',all.x=F)
```

```{r, echo=FALSE}
#Create the map
library(RColorBrewer)
library(classInt)
varToPLot=WA_ex_zip$amount

numberOfClasses = 5
colorForScale='YlGnBu'
colors = brewer.pal(numberOfClasses, colorForScale)
intervals <- classIntervals(varToPLot, numberOfClasses, 
                            style = "quantile",
                            dataPrecision=2)

colorPallette <- findColours(intervals, colors)

legendText="Expenditure amount"
shrinkLegend=0.4
title="Distibution of aggregated expenditure of WA in 2016"

plot(wazipMap,col='white',main=title)
plot(expenWAmap, col = colorPallette,border='grey',add=T)

legend('topright', legend = names(attr(colorPallette, "table")), 
       fill = attr(colorPallette, "palette"), cex = shrinkLegend, 
       bty = "n",
       title=legendText)
```

We could see that recipients from zipcode '98109', '98104', '98122', '98115', '98134', '98194', and '98101'received the higher amount of expenditures.
```{r, echo=FALSE}

WA_ex_zip$recipient_zip=as.character(WA_ex_zip$recipient_zip)
WA_ex_zip %>% 
  filter(
    recipient_zip %in% c('98109', '98104', '98122', '98115', '98134', '98194', '98101') 
    ) %>%
  ggplot(aes(x=recipient_zip, y=amount)) + 
  geom_bar(aes(fill = amount), position = "dodge", stat="identity")+
  labs(x = 'Recipient Zipcode', y = 'Total Expenditure', title='Zipcode with the highest Expenditure')

```
We could see the distribution of total expenditures in those areas, but there could have more than one recipient in one area. I hope to know how many recipients are there in each zipcode, and the following table shows the result. The area with zipcode 98032 has the most recipients, which is 306.
```{r, echo=FALSE}
Num_zip=WA2016[(WA2016$recipient_zip <=99403) & (WA2016$recipient_zip>=98001),]
Num_zip2 <- as.data.frame(table(Num_zip$recipient_zip))
colnames(Num_zip2)<- c("recipient_zip","Freq")
Num_zip2[rev(order(Num_zip2$Freq)),]
```

```{r, echo=FALSE}
#Merge the two data frame we got above
A=merge(WA_ex_zip,Num_zip2, by.x='recipient_zip', by.y='recipient_zip',all.x=F)
```
I got the total expenditure of per zipcode, and got the number of recipients in that area. I could easily get the average amount of expenditures under per zipcode.
```{r, echo=FALSE}
Average=A$amount/A$Freq
summary(Average)
```

```{r, echo=FALSE}
#Add Average column to dataframe A, name the new dataframe as B
B <-cbind(A,Average)

```
I plotted the average expenditure on the map too. We could compare the map with the last one, and could see that the locations in color bule has changed.
```{r, echo=FALSE}
AveWAmap=merge(wazipMap,A, by.x='ZCTA5CE10', by.y='recipient_zip',all.x=F)
varToPLot2=B$Average

numberOfClasses = 5
colorForScale='YlGnBu'
colors = brewer.pal(numberOfClasses, colorForScale)
intervals <- classIntervals(varToPLot2, numberOfClasses, 
                            style = "quantile",
                            dataPrecision=2)

colorPallette <- findColours(intervals, colors)

legendText="Expenditure amount"
shrinkLegend=0.4
title="Distibution of average expenditure of WA in 2016"

plot(wazipMap,col='white',main=title)
plot(AveWAmap, col = colorPallette,border='grey',add=T)

legend('topright', legend = names(attr(colorPallette, "table")), 
       fill = attr(colorPallette, "palette"), cex = shrinkLegend, 
       bty = "n",
       title=legendText)
```

In the end of the analysis, I also explored How different parties contribute to the expenditure for the place with the highest expenditure?
```{r, echo=FALSE}
WA2016$recipient_zip=as.character(WA2016$recipient_zip)
WA2016 %>% 
  filter(
    recipient_zip %in% c('98109', '98104', '98122', '98115', '98134') 
    ) %>%
  ggplot(aes(recipient_zip, amount)) + geom_bar(aes(fill = party), position = "dodge", stat="identity")
```

#####Which party has the highest government expenditure in WA in 2016?
```{r, echo=FALSE}
WA2016 %>% 
  ggplot(aes(party, fill = amount)) + 
  geom_bar() + 
  labs(
    x = 'Party'
    , y = 'Expenditure amount'
    , fill = element_blank()
    ) 
```


